---
title: "General Social Survey"
author: "Johanna Iniguez"
date: "10/7/2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey.  

The first question is this: is opinion on the death pentay (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$ : there is no  relationship between opinion on the death penalty and gun ownership.

$H_A$ : there is a relationship between opinion on the death pentaly is not independent of gun ownership.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables are categorical: hence we use bar graphs to graphically show the data, and a cross-tabulation for a numerical summary . A chi-squared test will be used to infer whether or not there is a relationship betweenn the varibles.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
 

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+OWNGUN, data=mydata4, main='Capital Punishment vs Gun Ownership', type ='percent', xlab ='Captial Punishment')
```

This bar chart shows a relationship between those who favor capital punishment are more likely to own a gun . While those who oppose capital punishment are less likley to own a gun .

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
tab1 = xtabs(~CAPPUN+OWNGUN, data=mydata4)
rowPerc(tab1)
```
This table shows a relationship in which those who favor capital punishment are more likely to own a gun . While those who oppose capital punishment are less likley to own a gun .

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisqtestGC(tab1)
```

The P-value for the chisq test is 0.02 < 0.05 so We reject the null hypothesis and conclude we have evidence to support the contention of gun ownership and whether one favors capital punishment are related. 


# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->


The second is this: Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)?

$H_0$ : There is no relationship between opinion on  the Big Bang theory and belief in the after life(POSTLIFE) .

$H_A$ :There is a relationship between opinion on the Big Bang Theory and belief in the after life(POSTLIFE) . 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables are categorical: hence we use bar graphs to graphically show the data, and a cross-tabulation for a numerical summary . A chi-squared test will be used to infer whether or not there is a relationship betweenn the varibles.



##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main='Big Bang Theory vs Post Life', type ='percent', xlab ='Big Bang Theory')

```

This bar chart shows a relationship between those who favor and oppose the Big Bang Theory also believe in PostLife. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(tab1)
```
This table shows a relation in opinion in which those who favor and oppose the Big Bang Theory also believe in PostLife.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```

The P value  for the chisqtest is 0.08< 0.05 so we fail to reject the null hypothesis and conclude we have evidence to support that there is no relationship between belief in the Big Bang Theory and the Afterlife. 

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The third question is : if a persons' opinion on death penalty (CAPPUN) is independent of their race (RACE).

$H_0$ : There is no relationship between opinion on  the death penalty (CAPPUN) and race .

$H_A$ :There is a relationship between opinion on the death penalty (CAPPUN) and race .


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables are categorical: hence we use bar graphs to graphically show the data, and a cross-tabulation for a numerical summary . A chi-squared test will be used to infer whether or not there is a relationship betweenn the varibles.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main='Death Penalty vs Race', type ='percent', xlab ='Death Penalty')

```

This bar chart shows that amongst white and other race there is a high factor in the death penalty while in the black race there is a high opposition towards the death penalty towards the total survey.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~CAPPUN+RACE, data=mydata4)
rowPerc(tab1)
```

This table shows a large favor of white and other support favoring capital punishment while the majority of black oppose capital punishment. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```

The p value for the chisq test is 0.0009<0.05 so we reject the null hypothesis that "there is no relationship between the death penalty and race" and conclude that there is a relationship between race and choice on favoring or opposing the death penalty. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The first and third question show strong relationship on the two variables as they both reject the null hypothesis and are questions aimed possibly towards social rules and settings. However, when looking at the question two regarding the Big Bang Theory and the belief in afterlife the data fails to reject the null hypothesis. This coulc be caused by the question which compares a science to a religion as the data shows how they are clearly independent as the two are considered seperate.  

